---
title: "Assignment ZK"
author: "Rachel Fischer"
date: "10/11/2020"
output: html_document
---

#LOAD PACKAGES
```{r}
if (!require(gridExtra)) install.packages("gridExtra") #plots
library(gridExtra)
if (!require(ggpubr)) install.packages("ggpubr") #plots
library(ggpubr)
if (!require(car)) install.packages("car") #residual plots
library(car)
if (!require(lmtest)) install.packages("lmtest") #bptest
library(lmtest)
if (!require(sandwich)) install.packages("sandwich") #coeftest
library(sandwich)
if (!require(tibble)) install.packages("tibble") #bootstrapping
library(tibble)
if (!require(psych)) install.packages("psych") 
library(psych)
if (!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
if (!require(kableExtra)) install.packages("kableExtra")
library(kableExtra)
if (!require(sjPlot)) install.packages("sjPlot")
library(sjPlot)
if (!require(lme4)) install.packages("lme4") #mixed model
library(lme4)
if (!require(cAIC4)) install.packages("cAIC4") #mixed model
library(cAIC4)
if (!require(insight)) install.packages("insight") #mixed model
library(insight)

coef_table = function(model) {
require(lm.beta)
mod_sum = summary(model)
mod_sum_p_values = as.character(round(mod_sum$coefficients[,
4], 3))
mod_sum_p_values[mod_sum_p_values != "0" & mod_sum_p_values !=
"1"] = substr(mod_sum_p_values[mod_sum_p_values != "0" &
mod_sum_p_values != "1"], 2, nchar(mod_sum_p_values[mod_sum_p_values !=
"0" & mod_sum_p_values != "1"]))
mod_sum_p_values[mod_sum_p_values == "0"] = "<.001"
mod_sum_table = cbind(as.data.frame(round(cbind(coef(model),
confint(model), c(0, lm.beta(model)$standardized.coefficients[c(2:length(model$coefficients))])),
2)), mod_sum_p_values)
names(mod_sum_table) = c("b", "95%CI lb", "95%CI ub", "Std.Beta",
"p-value")
mod_sum_table["(Intercept)", "Std.Beta"] = "0"
return(mod_sum_table)
}
```


#ASSIGNMENT 1
##load data
```{r}
Ass_1_raw_data <- read.csv("https://tinyurl.com/ha-dataset1")
Ass_1 <- Ass_1_raw_data
```

## data cleaning
```{r}
#checking errors
Ass_1 %>% summary()
hist(Ass_1$STAI_trait, main = "STAI_trait")

hist(Ass_1$age, main = "age")

hist(Ass_1$pain_cat, main = "pain cat")

hist(Ass_1$mindfulness, main = "mindfulness")



#correct errors recoding or exclusion
Ass_1$STAI_trait[Ass_1$STAI_trait==3.9]<- 39 #typing error
Ass_1$age[Ass_1$age==444]<- 44 #typing error
Ass_1$household_income[Ass_1$household_income==-3732]<- 3732 #typing error
Ass_1 <- subset(Ass_1, IQ != 49) # exclusion of participant with a really low IQ --> moderate mental retardation

#centering metric variables
# metric presictors are centered that it is easier to interpret the intercept and make the models easier to compare
Ass_1$age_c <- scale(Ass_1$age, center=T, scale = F)
Ass_1$STAI_trait_c <- scale(Ass_1$STAI_trait, center=T, scale = F)
Ass_1$pain_cat_c <- scale(Ass_1$pain_cat, center=T, scale = F)
Ass_1$serum_c <- scale(Ass_1$cortisol_serum, center=T, scale = F)
Ass_1$saliva_c <- scale(Ass_1$cortisol_saliva, center=T, scale = F)
Ass_1$mind_c <- scale(Ass_1$mindfulness, center=T, scale = F)
Ass_1$IQ_c <- scale(Ass_1$IQ, center=T, scale = F)
Ass_1$weight_c <- scale(Ass_1$weight, center=T, scale = F)
Ass_1$income_c <- scale(Ass_1$household_income, center=T, scale = F)

```


##analysis
###descriptives
```{r}
Ass_1 %>% 
  summarize(number = n(), mean_age = mean(age), sd_age = sd(age), mean_STAI = mean(STAI_trait), sd_STAI = sd(STAI_trait), mean_pain = mean(pain), sd_pain = sd(pain), mean_pain_cat = mean(pain_cat), sd_pain_cat = sd(pain_cat), mean_cor_se = mean(cortisol_serum), sd_cor_se = sd(cortisol_serum), mean_cor_sa = mean(cortisol_saliva), sd_cor_sa = sd(cortisol_saliva), mean_mind = mean(mindfulness), sd_mind = sd(mindfulness))

# by sex
Ass_1 %>% 
  group_by (sex) %>% 
  summarize(number = n(), mean_age = mean(age), sd_age = sd(age), mean_STAI = mean(STAI_trait), sd_STAI = sd(STAI_trait), mean_pain = mean(pain), sd_pain = sd(pain), mean_pain_cat = mean(pain_cat), sd_pain_cat = sd(pain_cat), mean_cor_se = mean(cortisol_serum), sd_cor_se = sd(cortisol_serum), mean_cor_sa = mean(cortisol_saliva), sd_cor_sa = sd(cortisol_saliva), mean_mind = mean(mindfulness), sd_mind = sd(mindfulness))
```

### models
```{r}
model1 <- pain ~ age_c + sex
model2 <-pain ~ age_c + sex + STAI_trait_c  + pain_cat_c + serum_c + saliva_c + mind_c
model2_a <-pain ~ age_c + sex + STAI_trait_c  + pain_cat_c + serum_c + mind_c

mod_1 <- lm(model1, data = Ass_1)
mod_2 <- lm(model2, data = Ass_1)

#adjusted models
reg_mod_1 <- lm(model1, data = Ass_1)
reg_mod_2 <- lm(model2_a, data = Ass_1)

```

###pre-tests

####Overview
```{r}
plot_age <- Ass_1 %>% ggplot() +aes(x = age, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_STAI <- Ass_1 %>% ggplot() +aes(x = STAI_trait, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_pain_cat <- Ass_1 %>% ggplot() +aes(x = pain_cat, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_serum <- Ass_1 %>% ggplot() +aes(x = cortisol_serum, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_saliva <- Ass_1 %>% ggplot() +aes(x = cortisol_saliva, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_cortisol <- Ass_1 %>% ggplot() +aes(x = cortisol_saliva, y = cortisol_serum,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1) # obviously multicolinearity
grid.arrange(plot_age, plot_STAI, plot_pain_cat, plot_serum, plot_saliva, plot_cortisol, nrow=3)
```
####multicollinearity
```{r}
mod_2 %>% vif() # bad VIFs for cortisol measurements --> exclusion of cortisol_saliva
vif(reg_mod_2)
```

####normality
```{r}
reg_mod_2 %>% plot(which=5)
reg_mod_2 %>% plot(which=4)
reg_mod_2 %>% plot(which=2)
res_mod_2 <- enframe(residuals(reg_mod_2))
res_mod_2 %>% ggplot() +aes(x = value) +geom_histogram()
describe(residuals(reg_mod_2))

```

####Linearity
```{r}
crPlots(reg_mod_2)
```
####Homoscedasticity
```{r}
reg_mod_2 %>% plot(which=3)
reg_mod_2 %>% ncvTest() # NCV test
reg_mod_2 %>% bptest() # Breush-Pagan test
```

### Hierarchical Regression
```{r}
#testing nested models
anova(reg_mod_1, reg_mod_2)
#centered values
tab_model(reg_mod_1, reg_mod_2, show.se = T, show.std = T, show.aic = T, dv.labels = c("model 1", "model 2"),
          pred.labels = c("(Intercept)", "age", "male", "State Trait Anxiety", "pain catastrophizing", "cortisol","mindfulness"),
          digits.p = 3, file = "regression_table_centered.doc", show.obs = F)

summary(reg_mod_2) # 0.4936
summary(reg_mod_1)
coef_table(reg_mod_2)
coef_table(reg_mod_1)
```
#ASSIGNMENT 2
##load data
```{r}
Ass_2 <- Ass_1
Ass_2_raw_data <- read.csv("https://tinyurl.com/ha-dataset2")
Ass_2_new <- Ass_2_raw_data

#centering metric variables
# metric presictors are centered that it is easier to interpret the intercept and make the models easier to compare
Ass_2_new$age_c <- scale(Ass_2_new$age, center=T, scale = F)
Ass_2_new$STAI_trait_c <- scale(Ass_2_new$STAI_trait, center=T, scale = F)
Ass_2_new$pain_cat_c <- scale(Ass_2_new$pain_cat, center=T, scale = F)
Ass_2_new$serum_c <- scale(Ass_2_new$cortisol_serum, center=T, scale = F)
Ass_2_new$saliva_c <- scale(Ass_2_new$cortisol_saliva, center=T, scale = F)
Ass_2_new$mind_c <- scale(Ass_2_new$mindfulness, center=T, scale = F)
Ass_2_new$IQ_c <- scale(Ass_2_new$IQ, center=T, scale = F)
Ass_2_new$weight_c <- scale(Ass_2_new$weight, center=T, scale = F)
Ass_2_new$income_c <- scale(Ass_2_new$household_income, center=T, scale = F)
```

##analysis
### descriptives
```{r}
Ass_2 %>% 
  summarize(number = n(), mean_age = mean(age), sd_age = sd(age), mean_STAI = mean(STAI_trait), sd_STAI = sd(STAI_trait), mean_pain = mean(pain), sd_pain = sd(pain), mean_pain_cat = mean(pain_cat), sd_pain_cat = sd(pain_cat), mean_cor_se = mean(cortisol_serum), sd_cor_se = sd(cortisol_serum), mean_cor_sa = mean(cortisol_saliva), sd_cor_sa = sd(cortisol_saliva), mean_mind = mean(mindfulness), sd_mind = sd(mindfulness), mean_income = mean(household_income),
            sd_income = sd(household_income), mean_IQ = mean(IQ), sd_IQ = sd(IQ),
            mean_weight = mean(weight), sd_weight = sd(weight))

# by sex
Ass_2 %>% 
  group_by (sex) %>% 
  summarize(number = n(), mean_age = mean(age), sd_age = sd(age), mean_STAI = mean(STAI_trait), sd_STAI = sd(STAI_trait), mean_pain = mean(pain), sd_pain = sd(pain), mean_pain_cat = mean(pain_cat), sd_pain_cat = sd(pain_cat), mean_cor_se = mean(cortisol_serum), sd_cor_se = sd(cortisol_serum), mean_cor_sa = mean(cortisol_saliva), sd_cor_sa = sd(cortisol_saliva), mean_mind = mean(mindfulness), sd_mind = sd(mindfulness), mean_income = mean(household_income),
            sd_income = sd(household_income), mean_IQ = mean(IQ), sd_IQ = sd(IQ),
            mean_weight = mean(weight), sd_weight = sd(weight))
```
###models
```{r}
#models
model3 <-pain ~ age_c + sex + STAI_trait_c  + pain_cat_c + serum_c  + mind_c + weight_c + IQ_c + income_c
model3_back <- pain ~ age_c + sex  + pain_cat_c + serum_c + mind_c + income_c



#diagnostic models
mod_3 <- lm(model3, data=Ass_2)

#backward regression
reg_mod_3 <- mod_3
reg_mod_3_back <- lm(model3_back, data=Ass_2)


#model assignment 1
theory_based_model <- reg_mod_2
```

###pre-tests
####Overview
```{r}
plot_IQ <- Ass_2 %>% ggplot() +aes(x = IQ, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_income <- Ass_2 %>% ggplot() +aes(x = household_income, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
plot_weight <- Ass_2 %>% ggplot() +aes(x = weight, y = pain,) +geom_point() +geom_smooth(method = "lm") + stat_cor(p.accuracy = 0.001, r.accuracy = 0.01, method = "pearson", label.x = 1)
grid.arrange(plot_age, plot_STAI, plot_pain_cat, plot_serum, plot_IQ, plot_weight,  plot_income, nrow=3)
```
####multicollinearity
```{r}
mod_3 %>% vif() # bad VIFs for cortisol measurements --> exclusion of cortisol_saliva
```

####normality
```{r}
mod_3 %>% plot(which=5)
mod_3 %>% plot(which=4)
mod_3 %>% plot(which=2)
res_mod_3 <- enframe(residuals(mod_3))
res_mod_3 %>% ggplot() +aes(x = value) +geom_histogram()
describe(residuals(mod_3))

```

####Linearity
```{r}
crPlots(mod_3)
```
####Homoscedasticity
```{r}
mod_3 %>% plot(which=3)
mod_3 %>% ncvTest() # NCV test
mod_3 %>% bptest() # Breush-Pagan test
```
### backward regression
```{r}
mod_back <- step(reg_mod_3, direction = "backward")
summary(mod_back)
AIC(mod_3)
anova(mod_back, mod_3)

tab_model(mod_back, theory_based_model, show.se = T, show.std = T, show.aic = T, dv.labels = c("backward model", "theory based model"), pred.labels = c("(Intercept)", "age", "sex [male]" ,"pain catastrophizing", "cortisol","mindfulness", "income", "State Trait Anxiety" ),digits.p = 3, file = "backward_regression_centered.doc", show.obs = F)

coef_table(mod_back)
```
### prediction
```{r}
#RSS differs centred - non centred values by 1
pre_mod_theory <- predict(reg_mod_2, Ass_2_new)
pre_mod_result <- predict(mod_back, Ass_2_new)

RSS_theory <- sum((Ass_2_new$pain - pre_mod_theory)^2) 
RSS_result <- sum((Ass_2_new$pain - pre_mod_result)^2) 
RSS_theory
RSS_result
```

#ASSIGNMENT 3
## load data
```{r}
Ass_3.1_raw_data <- read.csv("https://tinyurl.com/ha-dataset3")
Ass_3.2_raw_data <- read.csv("https://tinyurl.com/ha-dataset4")
Ass_3.1 <- Ass_3.1_raw_data
Ass_3.2 <- Ass_3.2_raw_data
```

##data cleaning
```{r}
Ass_3.1 %>% summary()
Ass_3.2 %>% summary()
Ass_3.2$mindfulness[Ass_3.2$mindfulness==6.05]<- NA
Ass_3.1$sex[Ass_3.1$sex=="femlae"]<- "female"
Ass_3.1$household_income[Ass_3.1$household_income==-6994]<- 6994 
Ass_3.1$household_income[Ass_3.1$household_income==-23482]<- 23482

#centering metric variables
# metric presictors are centered that it is easier to interpret the intercept and make the models easier to compare
Ass_3.1$age_c <- scale(Ass_3.1$age, center=T, scale = F)
Ass_3.1$STAI_trait_c <- scale(Ass_3.1$STAI_trait, center=T, scale = F)
Ass_3.1$pain_cat_c <- scale(Ass_3.1$pain_cat, center=T, scale = F)
Ass_3.1$serum_c <- scale(Ass_3.1$cortisol_serum, center=T, scale = F)
Ass_3.1$saliva_c <- scale(Ass_3.1$cortisol_saliva, center=T, scale = F)
Ass_3.1$mind_c <- scale(Ass_3.1$mindfulness, center=T, scale = F)
Ass_3.1$IQ_c <- scale(Ass_3.1$IQ, center=T, scale = F)
Ass_3.1$weight_c <- scale(Ass_3.1$weight, center=T, scale = F)
Ass_3.1$income_c <- scale(Ass_3.1$household_income, center=T, scale = F)

Ass_3.2$age_c <- scale(Ass_3.2$age, center=T, scale = F)
Ass_3.2$STAI_trait_c <- scale(Ass_3.2$STAI_trait, center=T, scale = F)
Ass_3.2$pain_cat_c <- scale(Ass_3.2$pain_cat, center=T, scale = F)
Ass_3.2$serum_c <- scale(Ass_3.2$cortisol_serum, center=T, scale = F)
Ass_3.2$saliva_c <- scale(Ass_3.2$cortisol_saliva, center=T, scale = F)
Ass_3.2$mind_c <- scale(Ass_3.2$mindfulness, center=T, scale = F)
Ass_3.2$IQ_c <- scale(Ass_3.2$IQ, center=T, scale = F)
Ass_3.2$weight_c <- scale(Ass_3.2$weight, center=T, scale = F)
Ass_3.2$income_c <- scale(Ass_3.2$household_income, center=T, scale = F)

Ass_3.1 %>%
mutate(hospital = factor(hospital))
Ass_3.2 %>%
mutate(hospital = factor(hospital))
```

##analysis

###models
```{r}
#dataset 3.1
mixed <- pain ~ age_c + sex + STAI_trait_c  + pain_cat_c + serum_c + mind_c + (1|hospital)
mixed_2_sl <- pain ~ serum_c + (serum_c|hospital)
mixed_2_int <- pain ~ serum_c + (1|hospital)

reg_mixed_3.1 <- lmer(mixed, data = Ass_3.1)
reg_mixed_3_sl <- lmer(mixed_2_sl, data = Ass_3.1)
reg_mixed_3_int <- lmer(mixed_2_int, data = Ass_3.1)
```

###pre-test
```{r}
#↨testing nomralitiy of the residuals to test if the sum of squares are influenced by it
reg_mixed_3.1 %>% plot(which=5)
reg_mixed_3.1 %>% plot(which=4)
reg_mixed_3.1 %>% plot(which=2)
reg_mixed_3.1 <- enframe(residuals(reg_mixed_3.1))
reg_mixed_3.1 %>% ggplot() +aes(x = value) +geom_histogram()
reg_mixed_3.1 <- lmer(mixed, data = Ass_3.1)
res <- residuals(reg_mixed_3.1)
describe(residuals(reg_mixed_3.1))
```

### mixed model
```{r}
tab_model(reg_mixed_3.1, reg_mod_2, show.se = T, show.std = T, dv.labels = c("mixed", "Ass 1"),pred.labels = c("(Intercept)", "age", "sex [male]" ,"State Trait Anxiety" ,"pain catastrophizing", "cortisol","mindfulness"), digits.p = 4, show.obs = F, file = "mixed_regression_centered.doc") #adjust AIC after running script!

stdCoef.merMod(reg_mixed_3.1)
summary(reg_mixed_3.1)
cAIC(reg_mixed_3.1)$caic
AIC(reg_mod_2)

pre_mod_3.1 <- predict(reg_mixed_3.1, Ass_3.2, allow.new.level=T)

RSS_mod_3.1 <- sum((Ass_3.2$pain - pre_mod_3.1)^2, na.rm = T)
TSS_mod_3.1 <- sum((Ass_3.2$pain - mean(Ass_3.2$pain))^2, na.rm = T)
1-(RSS_mod_3.1/TSS_mod_3.1)

pre_mod_old <- predict(reg_mixed_3.1, allow.new.level=T)

RSS_old <- sum((Ass_3.1$pain - pre_mod_old)^2, na.rm = T)
TSS_old <- sum((Ass_3.1$pain - mean(Ass_3.1$pain))^2, na.rm = T)
1-(RSS_old/TSS_old)

mean(Ass_3.1$age)
get_variance(reg_mixed_3.1)
```

###slope
```{r}
Ass_3.1 <- Ass_3.1 %>%
mutate(pred_int = predict(reg_mixed_3_int))

Ass_3.1 %>%
ggplot() +
aes(y = pain, x = cortisol_serum, group = hospital)+
geom_point(aes(color = hospital), size = 4) +
geom_line(color='red', aes(y=pred_int, x=cortisol_serum))+
facet_wrap( ~ hospital, ncol = 2)+
xlab("cortisol")+
ggsave("slope_int.jpg")

```